---
title: "Assignment 7"
author: "Emily MacQuarrie"
date: "3/1/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r libraries}
library(tseries)
library(forecast)
library(vars)
```

## Question 1
```{r data}
load("./motel.rda")
motel <- as.data.frame(motel)
# calculate cost
motel$Cost <- motel$Takings * 1000 / motel$Roomnights
cpi <- ts(motel$CPI, start=c(1980,1), end=c(1995, 6), frequency=12)
cost <- ts(motel$Cost, start=c(1980,1), end=c(1995, 6), frequency=12)
# part a: plot data
autoplot(cpi)
autoplot(cost)
# part b: correlation
dev.new(width=5, height=4)
plot(cpi, cost)
# pearson correlation coefficient between CPI and Cost
cor(cpi, cost)
# part c: log transform
log_cpi <- log10(cpi)
log_cost <- log10(cost)

autoplot(log_cpi)
autoplot(log_cost)
cor(log_cpi, log_cost)
```
The Pearson correlation coefficient between CPI and Cost is 0.991, which tells us that the values are very highly correlated with each other. Based on the CPI and Cost plots, a logarithmic transformation appears to be appropriate here since the data is not linear and looks to decrease rate of increase as time progresses. Furthermore, the variance increases over time in these plots,especially in the Cost graph, and a logarithmic transformation could reduce this variation to become more stationary.

After taking the log transform, the new Pearson correlation coefficient between log CPI and log Cost is 0.993, which is marginally higher than the untransformed data.

## Question 2
```{r linear model}
# part a: linear model with log transformation
loglm <- tslm(log_cost ~ log_cpi)
# part b: plot relationship
plot(log_cost ~ log_cpi)
abline(loglm)
# part c: summarize model
summary(loglm)
checkresiduals(loglm)
```

According to the model results, the coefficient for the log of CPI is significant with a value of 1.303. However, the residuals appear to have a cyclical quality to them, and the ACF decreases relatively slowly. Therefore, the residuals do not appear to be stationary. The Breusch-Godfrey test also indicates that there exists some significant autocorrelation in the residuals.

## Question 3
```{r arima}
ar1 <- auto.arima(cost, xreg=cpi, lambda=0, D=1)
# part a: summarize model
summary(ar1)
# part b: residuals
checkresiduals(ar1)
``` 
Results of the auto.arima function show us that the best ARIMA model for this data is Regression with ARIMA(0,1,1)(0,1,2)[12] errors. The AICc is 863.79, and the BIC is -848.38.

The Ljung-Box test on the model has a p-value of 0.3, so we accept the null hypothesis. This implies that the residuals are independently distributed. Furthermore, the residual plot appears to be stationary, and the ACF plot looks to have minimal autocorrelation, aside from lags 18 and 34. Overall, this model appears at first glance to be a good fit for the data.

## Question 4
```{r naive}
# part a: naive forecast on CPI
na <- naive(cpi, lambda=0, h=18)
na
plot(na)
# part b: average price per room
arfc <- forecast(ar1, xreg=na$mean, h=18)
arfc
plot(arfc)

```
The naive forecast for CPI with a log transformation is 116.2 for the next 18 months. The result of the forecast in the Regression with ARIMA errors model shows that the data will vary with little to no positive trend over the next 18 months, ranging from 81 to 98 dollars per room. The price is expected to be the lowest in April 1996 and highest in November 1995 within those 18 months.

## Question 5
```{r var}
# part a: var model
vars <- cbind(log_cost, log_cpi)
var <- VAR(vars, p = 10, type="both", season=12)
summary(var)
# part b: forecast
varfc <- predict(var, n.ahead=18)
varfc
plot(varfc, names="log_cost")
plot(varfc, names="log_cpi")
# part c: residuals
acf(resid(var))
```
According to the results of the model, the log CPI appears to be positively correlated with the lag 1 log CPI variable, the lag 2 log Cost variable, and the lag 3 log Cost variable. The log Cost appears to be positively correlated with the lag 1 log CPI variable, the lag 6 log CPI variable, and the lag 9 log CPI variable. The log Cost appears to be negatively correlated with the lag 8 log Cost and the lag 10 log CPI variables. The forecasts for both variables appear to be increasing with time.

The log cost residuals in the upper right graph show that autocorrelation is not very significant, except for lag 18, which is on the edge of being significant. There is more autocorrelation in the log of cost than the log of CPI which can be seen in the bottom right graph. This indicates that residuals do not exhibit autocorrelation, and the lack of autocorrelation is especially strong in the log of CPI. Furthermore, the log CPI and log Cost do not appear to be significantly autocorrelated with each other. The lack of autocorrelation in the residuals indicates that the model is a relatively good fit for the data.